---
author: "Kevin McNulty"
title: "R Notebook"
output: html_notebook
date: "Dec 8 2022"
---
##About this notebook
Created by Kevin McNulty

1: Where are your questions that you are trying to answer. they need to be on this document.
2: What is the data, where does it come from and what do the variables mean. We discussed using this: https://collegescorecard.ed.gov/data/documentation/ That is a major gap.
3: Filter the college scorecard for all maryland schools and find the max, min, mean and median for student load debt. graph that. 
4: What school has the highest debt per capita? do that math, get your top 10 and bottom 10 schools and graph that.
5: I don't know what you get by joining to College Navigator except it filters the maryland schools to four year schools. the only useful metric is the graduation rate - the rest you have on the college scorecard data. So if you do this, I would just be focusing on the college scorecard data.

I'm conducting data analysis of student loan debt for college graduates in the state of Maryland. 

## Load libraries and establish settings
```{r}
options(scipen=999)
library(tidyverse)
library(janitor)
library(ggplot2)
```
##Load data



```{r}
college_scorecard <- read.csv("KEVIN_MERGED2019_20_PP_DEC6.csv") %>% 
   clean_names()
maryland_schools <- read.csv("CollegeNavigator_Search_2022-11-17_13.47.02.csv") %>% 
  clean_names()
```

##Filter for Maryland 


```{r}
college_scorecard %>% 
  filter(stabbr =="MD")
```

##Join both dataframes 


```{r}
  colnames(maryland_schools)[colnames(maryland_schools) == "ipeds_id"] ="unitid"

  college_scorecard <- maryland_schools %>%
  inner_join(college_scorecard, by="unitid")
```

##Clean columns
```{r}
  college_scorecard <- college_scorecard [-c(2, 3, 7, 10, 11, 12, 14, 16, 17, 19)]
  colnames(college_scorecard)[colnames(college_scorecard) == "npt4_pub"] ="avg_net_price"
  colnames(college_scorecard)[colnames(college_scorecard) == "stabbr"] ="state"
  colnames(college_scorecard)[colnames(college_scorecard) == "grad_debt_mdn"] ="median_debt"
  colnames(college_scorecard)[colnames(college_scorecard) == "ugds"] ="undergrads"
```

##Make columns numeric; remove null values
```{r}
college_scorecard$median_debt <- gsub("PrivacySuppressed", "", college_scorecard$median_debt)
college_scorecard$median_debt <- as.numeric(college_scorecard$median_debt)
```

##Q1 Graduates of which Maryland schools are in the most debt, on average?
  
##Question 2